---
title: "Let's Talk About How To Ease Re-Analysis"
author: "Jake Bowers"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  beamer_presentation:
    keep_tex: yes
    latex_engine: xelatex
    slide_level: 2
    template: sbst.beamer
    toc: yes
  ioslides_presentation:
    css: sbstpresentation.css
    logo: ../graphics/SBSTLightBulb.png
    smaller: yes
---
```{r setup, include=FALSE, echo=FALSE, cache=FALSE}
## the 'toc: yes' line above tricks pandoc into running latex twice so that we can see the nice logos

# Some customization.  You can alter or delete as desired (if you know what you are doing).
# knitr settings to control how R chunks work.

require(knitr)
opts_chunk$set(
  size = "scriptsize",    # slightly smaller font for code
  echo = TRUE,
  results = 'markup',
  strip.white = TRUE,
  cache = FALSE,
  highlight = TRUE,
  width.cutoff = 60,
  out.width = '.9\\textwidth',
  message = FALSE,
  comment = NA,
  tidy = TRUE,     # display code as typed
  tidy.opts = list(blank = FALSE, width.cutoff = 60))
```

# Overview: The Code Problem

## The Code Problem

Data analysis is computer programming.

\medskip

Computer programs have bugs.

\medskip

How can anyone do high stakes data analysis?  

\smallskip

(Ideas for how to do this on the next slide.)


## General Strategies to Enhance Confidence and Reduce Bugs
<!-- HTML presentations use a different color choice method  for example <span style="color:orange">Work together</span> --->


1) \textcolor{orange}{Work together:} 
	- If you and your colleagues are going to have to use each others' code, you may start writing code as if others were going to read it.
	- A second pair of eyes spots errors.

2) \textcolor{orange}{Test assumptions:} Can we get the program to fail when it does something wrong?

3) \textcolor{orange}{Track dependencies:} What file depends on what other file? Can we go from raw data to results in one command? Where did this figure come from?
	- Share files instead of copying files.
	- Think in Batch mode (even if you develop in interactive mode)
	- Plan for use of code on a fresh computer/empty directory.
  
4) Other ideas for after the presentation? 

# Track dependencies: Know how outputs relate to inputs

## Data Sharing Idea 1: Google Sheets

```{r googlesheets1, size="tiny", tidy=TRUE}
## The first time this is run in a while, you'll need to run each line by hand. Thereafter, batch mode will be fine
library(googlesheets)
fakedatGS <- gs_title("experimentDat.csv") ## get info about the data
fakedat <- as.data.frame(gs_read_csv(fakedatGS,as.is=TRUE),stringsAsFactors=FALSE)
str(fakedat)
```

```{r genexample, echo=FALSE,results='hide'}
fakedat$y0 <- rnorm(nrow(fakedat))+5 ## no negative numbers
fakedat$y1 <- fakedat$y0+5
fakedat$Y <- with(fakedat, Z * y1 + (1-Z) * y0)
```

## Data Sharing Idea 2: Data Online

Other data can be read directly from a website:
```{r}
usaCountiesPopCentroids<-read.csv(url("http://www2.census.gov/geo/docs/reference/cenpop2010/county/CenPop2010_Mean_CO.txt"),as.is=TRUE)
head(usaCountiesPopCentroids)
```

## Other Data Sharing Ideas

Small Text Files (.txt,.csv) can live on Github (currently this is all public)

\medskip

Eventually we will have some shared data space on an Amazon EC2 instance.

## Think in Batch Mode

\textcolor{orange}{Ideal:} Go from raw data to numbers, tables, figures in one command starting from a clean slate.

For example, at the command line `R --vanilla --file=masterfile.R` does the whole analysis.

```{r masterfileexample, eval=FALSE, size="tiny"}
## Master file for Project X

source("data/datacleanup.R") ## produces wrkdat.rda
rm(list=ls()) ## delete everything
source("data/randcheck.R") ## uses wrkdat.rda and produces randok.rda
rm(list=ls())
source("data/analysis.R") ## does the analysis
rm(list=ls())
source("data/figure1.R") ## makes figures/fig1.pdf
rm(list=ls())
source("data/table1.R") ## makes figures/table1.pdf
```


## Another Example: Literate Programming via Rmarkdown

This presentation is an example of a one-file approach where the figures and tables and numbers (like the size of the dataset `r nrow(fakedat)`) are generated on the fly so that we know that the figures and tables and numbers were created within the file.


```{r map, echo=FALSE, fig.show='asis',fig.keep='high', fig.cap="Dark counties assigned to some treatment. Light counties assigned to no treatment",out.width=".7\\textwidth"}
par(oma=rep(0,4))
library(maps)
data(countyMapEnv)
data(usaMapEnv)
data(county.fips)
data("state.fips")
map('state',regions=state.fips$polyname[state.fips$abb %in% fakedat$state],exact=FALSE,mar=c(1,1,1,1))
map("county",regions=county.fips$polyname[county.fips$fips %in% as.numeric(fakedat$county)],add=TRUE)
map("county",regions=county.fips$polyname[county.fips$fips %in% as.numeric(fakedat$county)[fakedat$Z==1]],col="gray",fill=TRUE,add=TRUE)
```

## Another example of literate programming

```{r xtab, results='asis'}
library(xtable)
resmat<-sapply(fakedat[,c("n","Z","Y")],function(x){ c(mean=mean(x),sd=sd(x))})
tab1<-xtable(resmat,caption = paste("Descriptives. Total N=",nrow(fakedat),".",sep=""))
print(tab1,comment=FALSE)
```



## For more complex projects: Makefiles

Here is a little Makefile for Project X. I would type `make memo.pdf` to run it all (or the pieces that haven't been done yet).

```
memo.pdf: figures/figure1.pdf reports/memo.Rmd
	Rscript -e 'library(rmarkdown); render('memo.Rmd')'
	
data/wrkdat.rda: data/datacleanup.R
	R --vanilla --file=data/datacleanup.R

figures/figure1.pdf: data/wrkdat.rda figures/figure1.R
	R --vanilla --file=figures/figure1.R
```

# Work Together: How to be nice to your future self and future friends with code?

## Interventions, Blocks, Outcomes!

Experiments are defined by \textcolor{orange}{treatment assignments}, fixed/pre-treatment \textcolor{orange}{design features} (like blocks/strata), and \textcolor{orange}{outcomes}.^[And we assume that, within block, the probability of treatment assignment is constant for each unit.]

\medskip

**Ideal:** Provide raw data+code and/or a rectangular dataset with those features as input to results production. For example:

```{r, echo=FALSE,results="asis"}
mat<-fakedat[c(2:3,9:12),-c(7,8)]
print(xtable(mat,
			 caption="Example of working data. Treatment ($Z=1$) assigned within pairs of counties (pair) within state. Outcome is Y."),comment=FALSE,include.rownames=FALSE)
```

## Sharing Code

Sharing code is better than copying code. 

\medskip

We already share docs with Google Drive. Github is better for code. We can talk more about Github later.

\medskip

\textcolor{orange}{Challenges of code sharing:}  Avoiding clobbering. Avoiding clutter. Fixing mistakes. Versioning (branching, forking).



# Unit Testing: Writing code is writing tests

## Create productive errors

For example, here I randomly assign treatment within pair of counties. If the code worked (and the preceding code setting up the counties and pairs worked), I should have exactly one treated per pair.

```{r exampletest, eval=FALSE, size='tiny'}
set.seed(20150101) ## to enable reproducible randomization
fakedat$Z<-unsplit(lapply(split(fakedat$county,fakedat$pair),
						  function(x){ sample(c(0,1)) 
						  	}),fakedat$pair)
head(fakedat) ## look at data

stopifnot(all(tapply(fakedat$Z,fakedat$pair,sum)==1)) ## exactly 1 treated county per pair
```

## Another example:

A quote from the Microloans 2.0 file

> Anchorage is still duplicated, so I removed it by hand. Eventually the input > file should only have one row per county.

```{r eval=FALSE, size="tiny"}
countyCounts<-countyCounts[!(countyCounts$n==4 & countyCounts$combifips=="02020"),]
stopifnot(all(table(countyCounts$combifips)==1))
```

# Conversation


